Practical Hilbert space approximate Bayesian Gaussian processes for probabilistic programming

### 2022-03-09 Paper/Manuscript_Stats&Comp/reply_letter_STCO.tex
Minor edits to R1 comments in reply letter

We have derived numerical equations that gather the relationships between the key parameters of the model (initially presented in form of a graph in Figure 6) which allow updating the parameters and checking whether the approximation is accurate (Section 4.3.1). We have also added a general and robust iterative procedure to derive minimally sufficient hyperparameter values that ensure good GP approximations via our HSGP approach (Sections 4.5 and 4.6).

@@

We have implemented the HSGP model and the diagnostic tool for four different kernels, the squared exponential, Matern-3/2, Matern-5/2, and periodic squared exponential kernels, which cover a broad range of learning functions in practical applications (Sections 4.3 and 4.4). Kernel combinations that satisfy the requirements can also be used. We have also given recommendations to update $m$ and $c$ and perform diagnostics on any other different stationary kernel from those implemented in our study, by analyzing the constants characterizing the relationships for each covariance function (Sections 4.2 and 4.7). Explanations on how to estimate the number of basis functions needed for an accurate approximation for any other covariance function have also been provided (Section 4.7).

In Section 3, we describe the formulation and implementation of the HSGP model in detail, a description that can be used to implement it for any other stationary kernel if required. The only kernel-specific function that the user needs to know is the Fourier transform of the kernel to obtain its spectral density (Section 3 and Section 4.7).

@@

Our experience tells us that the range of expertise in users of probabilistics programming languages such as Stan and their higher level interfaces such a brms is quite wide. We don't want to speculate on the percentage of users that would be able to come up with comparable solutions themselves. To study this empirically, one had to conduct extensive user studies and their are not in scope of this paper.
We want to highlight though that rerunning of models and deciding about good updating rules takes a lot of human and compute time. We substantially reduce that time by reducing the number of model iterations, the computational requirements for each iteration, as well as eliminating the time an educated practioners would have had to spend to come up with some sensible rule themselves.

@@

\item Case study IV (Leukemia data set) is a 4D function with a multilevel GP structure. This case study is also a good example that by using our diagnostic tool you can easily get an optimal solution in 4D which makes it significantly faster than regular GPs. Otherwise, user would had probably been struggling searching for useful values for the hyperparameters without any guidance.
\end{itemize}

@@

The HSGP model is definitely useful in 1D and 2D, even for very wiggly functions, unless the number of data points is very small ($n \lesssim 300$) where the regular formulation of a GP is already reasonably fast itself.
It is difficult to give clear guidelines of when the application of the HSGP model can result to be impractical in other multidimensional cases, basically because it ultimately depends on the wigglyness of the function to be learned which is, in most of the cases, unknown. For that reason, in fact, the use of our diagnostics tool can be essential because it allows to iteratively fit the model using minimum computational requirements in each iteration. Observing how the requirements change in each iteration can be useful for the user to be able to recognize when the approximation is unfeasible to be applied. For $D>5$, even for very smooth functions, HSGPS definitely seem to be impractical. For $D<3$ it is widely useful and significantly faster than the exact GP even for very wiggly functions.

We disagree with the reviewer that the HSGP performs poorer than the exact GP on this Leukemia data set. In fact using the diagnostic tool the user will discover easily the optimal values for $c$ and $m$ that allows for a significantly faster performance than using the exact GP. Without the guide of the diagnostic tool, the user would have likely made several trials until finding an accurate and practical solution, consuming a lot of computation requirements in this highly demanding 4D case. Insofar, the understanding that "its applicability to multidimensional cases is mostly negative" is not fully accurate, as it ultimately depends on the wigglyness of the function. We now better discuss this point in various places in the paper. Specifically, in Section 4.8, we provide details on the performance of the model in multi-dimensional cases.

@@

Thank you again for your thorough review. To us, also empirical analysis provides scientific insights, in particular as we can study much more complicated scenarios that we cannot feasibly approach my mathematical analysis alone. As such, we think our paper nicely complements existing research.

Solin and Sarka (2020) fundamentally focus on the mathematical development of the model, while our work is necessary to facilitate the easier use of the method for practical users. We also analyze how the key factors of the model relate each other for accurate approximations. We derive numerical functions for these relationships that allow to update the parameters, perform diagnostics, and provide clear and useful guidelines for the whole procedure. So, we make it manageable for practical users, otherwise the practical applicability of the method would be highly limited.

>>>>

We have derived numerical equations that gather the relationships between the key parameters of the model (initially presented in form of a graph in Figure 6) which allow updating the parameters and checking whether the approximation is accurate (Section 4.3.1). We have also added a general and robust iterative procedure to derive minimally sufficient hyperparameter values that ensure accurate GP approximations via our HSGP approach (Sections 4.5 and 4.6).

@@

We have implemented the HSGP model and the diagnostic tool for four different kernels, the squared exponential, Matern-3/2, Matern-5/2, and periodic squared exponential kernels, which cover a broad range of learning functions in practical applications (Sections 4.3 and 4.4). Kernel combinations that satisfy the requirements can also be used. We have also given recommendations to update $m$ and $c$ and perform diagnostics on any other different stationary kernel from those implemented in our study, by analyzing the constants characterizing the relationships for each covariance function (Sections 4.2 and 4.7).

In Section 3, we describe the formulation and implementation of the HSGP model in detail, a description that can be used to implement it for any other stationary kernel if required. The only kernel-specific function that the user needs to know is the Fourier transform of the kernel to obtain its spectral density (Section 3). Moreover, we discuss how to estimate the number of basis functions needed for an accurate approximation for any other (stationary) covariance function have also been provided (Section 4.7).

@@

Our experience tells us that the range of expertise in users of probabilistics programming languages such as Stan and their higher level interfaces such a brms is quite wide. We don't want to speculate on the percentage of users that would be able to come up with comparable solutions themselves. To study this empirically, one had to conduct extensive user studies and their are not in scope of this paper.
We want to highlight, though, that rerunning models iteratively for estimating reasonable hyperparameter values often takes a lot of human and compute time. We substantially reduce that time by reducing the number of model iterations, the computational requirements for each iteration, as well as eliminating the time an educated practioners would have had to spend to come up with some sensible rule themselves.

@@

\item Case study IV (Leukemia data set) is a 4D function with a multilevel GP structure. This case study is also a good example that by using our diagnostic tool you can easily get an optimal solution in 4D which makes it significantly faster than regular GPs.
% Otherwise, the user would had probably been struggling searching for useful values for the hyperparameters without any guidance.
\end{itemize}

@@

The HSGP model is definitely useful in 1D and 2D, even for very wiggly functions, unless the number of data points is very small ($n \lesssim 300$) where the regular formulation of a GP is already reasonably fast itself.
It is difficult to give specific guidelines for when the HSGP model will be impractical for multidimensional problems, because it ultimately depends on the wigglyness/ of the function to be learned which is, in most of the cases, unknown. For that reason, in fact, the use of our diagnostics tool can be essential because it allows to iteratively fit the model using minimum computational requirements in each iteration. Observing how the requirements change in each iteration can be useful for the user to be able to recognize when the approximation is unfeasible to be applied. For $D>5$, even for very smooth functions, HSGPs definitely seem to be impractical. For $D<3$ it is widely useful and significantly faster than the exact GP even for very wiggly functions.

We disagree with the reviewer that the HSGP performs worse than the exact GP on this Leukemia data set. In fact, using the diagnostic tool the user will easily discover the optimal values for $c$ and $m$ that implies significantly faster computation than using the exact GP. Without the guide of the diagnostic tool, the user could easily end up using several trials until finding an accurate and practical solution, consuming a lot of computation requirements in this highly demanding 4D case. Insofar, the understanding that "its applicability to multidimensional cases is mostly negative" is not fully accurate, as it ultimately depends on the wigglyness of the function. We now better discuss this point in various places in the paper. Specifically, in Section 4.8, we provide details on the performance of the model in multi-dimensional cases.

@@

Thank you again for your thorough review. To us, empirical analysis also provides scientific insights, in particular as we can study much more complex scenarios that we cannot feasibly approach via mathematical analysis alone. As such, we think our paper nicely complements existing research.

Solin and Sarka (2020) fundamentally focus on the mathematical development of the model, while our work is necessary to decribe and facilitate applied use of the approximation. We also analyze how the key factors of the model relate to each other for accurate approximations. We derive numerical functions for these relationships that allow to update the parameters, perform diagnostics, and provide clear and useful guidelines for the whole procedure. So, we make it manageable for practical users, otherwise the practical applicability of the method would be highly limited.
